---
title: "Newsletter Helpers"
format:
  html:
    embed-resources: true
---

```{r}
library(dplyr)
library(readr)
library(here)
library(janitor)
library(lubridate)
library(stringr)
library(tidyr)
library(yaml)
library(purrr)
library(jsonlite)
library(tibble)
library(fs)
library(DT2)
```

This notebook contains helper snippets for the US-RSE newsletter. It
is intended to parse data into YAML before it is used in the newsletter
template.

The directory should be self-contained, so that it doesn't interfere
with any other US-RSE Website content. No software is installed,
and no changes are being made to the US-RSE Website. The notebook is 
intended to be run in a local environment, not as part of CI/CD.

## Old Data

Let's read in the old data to rbind it:

```{r}
old_data <- read_yaml(here("_data/newsletter-events-opportunities.yml")) |>
    map_dfr(as_tibble) |>
    mutate(
        expires = as_date(expires, format = "%Y-%m-%d")
    )
```

## Slack Data

Next, we'll grab the Slack data from the Newsletter Content Triage
[List](https://usrse.slack.com/lists/T8ZT4PJSW/F08MLHYKALR). 
We could use the Slack API, but I think it's actually easier
to export the list content as CSV and `git` track it in the 
`_data` directory.

Importantly, the most recent file is the one we want to keep;
Slack doesn't automatically overwrite the CSV, so we have to manually delete old files.

```{r}
old_slack_files <- list.files(here("_data/newsletter"), pattern = "Newsletter_Content_Triage", full.names = TRUE) |>
    file.info() |>
    rownames_to_column("file") |>
    as_tibble() |>
    arrange(desc(mtime))

if(nrow(old_slack_files) > 1){
    slack_file <- old_slack_files |>
        slice(1)

    walk(old_slack_files$file[-1], fs::file_delete)
    fs::file_move(slack_file$file, here("_data/newsletter/Newsletter_Content_Triage.csv"))
} else {
    message("No old files to delete.")
}
```

```{r}

slack <- read_csv(here("_data/newsletter/Newsletter_Content_Triage.csv")) |>
    clean_names()

section_col <- str_subset(names(slack), "what_section")

slack <- slack |>
    separate_rows(all_of(section_col), sep = ",\\s*") |>
    mutate(
        submission_date = as_date(submission_date, format = "%m/%d/%y, %I:%M %p"),
        staging = as.factor(staging)
        )    |>
    rename(section = all_of(section_col))
```

Now get the new events:

```{r}

fromJson_vec <- Vectorize(function(x){

    tryCatch({
        dat <- fromJSON(x)

        display <- dat$displayName
        url <- dat$originalUrl

        str_c("[", display, "](", url, ")")
    }, error = function(e){
        x
    })
}, USE.NAMES = FALSE)

slack |>
    filter(str_detect(section, "Event")) |>
    filter(staging == "Approved" | staging == "Published") |>
    mutate(
        content_expiration_date = case_when(
            is.na(content_expiration_date) ~ as_date("2026-08-01"),
            TRUE ~ content_expiration_date
        ),
        event_date = case_when(
            is.na(event_date) ~ as_date("2026-08-01"),
            TRUE ~ event_date
        ),
        type = case_when(
            str_detect(str_to_lower(content), "submit|application|submission|deadline") ~ "opportunity",
            TRUE ~ "event"
        ),
        location = case_when(
            !is.na(location) ~ location,
            str_detect(str_to_lower(content), "online|zoom|webinar|virtual") ~ "Online",
            TRUE ~ "In Person"
        ),
        link = fromJson_vec(link)
    ) -> events_opportunities
```

## Binding

Now we just put em together: 

```{r}
events_opportunities |>
    mutate(event_date = as.character(event_date)) |>
    select(
        expires = content_expiration_date,
        type = type,
        title = submission,
        when = event_date,
        where = location,
        preamble = content,
        links = link
    ) |>
    replace_na(list(when = as_date("2026-01-01"), where = "TBD", links = "TBD")) |>
    bind_rows(old_data) -> all_events_opportunities
```

Number of rows to publish next: 

```{r}
all_events_opportunities |>
    filter(expires >= Sys.Date()) |>
    dt2(compact = TRUE, striped = TRUE, hover = TRUE, font_scale = 0.85,
    options = list(pageLength = 10))
```

We can write this back to yaml now:

```{r}
all_events_opportunities |>
    filter(expires >= Sys.Date()) |>
    mutate(when = as.character(when)) |>
    mutate(expires = as.character(expires)) |>
    mutate(index= row_number()) |>
    group_by(index) |>
    group_split() |>
    write_yaml(here("_data/newsletter/events-opportunities.yml"))
```


