---
title: "Newsletter Helpers"
format:
  html:
    embed-resources: true
---

```{r}
library(dplyr)
library(readr)
library(here)
library(janitor)
library(lubridate)
library(stringr)
library(tidyr)
library(yaml)
library(purrr)
library(jsonlite)
library(tibble)
library(fs)
library(DT2)
```

This notebook contains helper snippets for the US-RSE newsletter. It
parses working inputs into YAML before they are used in the newsletter
template.

The edition is prepared locally. Running the notebook can update the shared
newsletter data under `_data`, but it does not deploy the website or install
software. Review those staged changes before handing an edition to the parent
site; this notebook is intended to be run locally, not as part of CI/CD.

## Old Data

Let's read in the old data to rbind it:

```{r}
# Slack exports use colon-delimited emoji names. Quarto can render literal
# Unicode emoji, but the final site parses the Liquid-generated content with
# Kramdown, which does not interpret Slack's :name: syntax. Convert the
# standard names used by this newsletter before the data reaches YAML.
slack_emoji_map <- c(
    happy = "🙂",
    smile = "😄",
    smiley = "😃",
    mega = "📣",
    seedling = "🌱",
    yellow_heart = "💛",
    point_right = "👉",
    tada = "🎉",
    shirt = "👕",
    unicorn_face = "🦄",
    trumpet = "🎺",
    bangbang = "‼️",
    speaking_head_in_silhouette = "🗣️",
    hotel = "🏨",
    slack = "💬",
    world_map = "🗺️",
    round_pushpin = "📍",
    eyes = "👀",
    point_down = "👇",
    clock4 = "🕓",
    dancer = "💃"
)

# These are Slack custom emoji rather than Unicode emoji. Use readable
# fallbacks so their names do not leak into the publication as raw tokens.
slack_custom_emoji_map <- c(
    usrse_main_logo = "US-RSE",
    spack = "Spack"
)

slack_emoji_to_unicode <- function(x) {
    replace_one <- function(token) {
        name <- str_remove_all(token, "^:|:$")

        if (name %in% names(slack_emoji_map)) {
            unname(slack_emoji_map[[name]])
        } else if (name %in% names(slack_custom_emoji_map)) {
            unname(slack_custom_emoji_map[[name]])
        } else {
            token
        }
    }

    str_replace_all(
        x,
        ":[A-Za-z0-9_+\\-]+:",
        function(tokens) vapply(tokens, replace_one, character(1))
    )
}

old_data <- read_yaml(here("_data/newsletter-events-opportunities.yml")) |>
    map_dfr(as_tibble) |>
    mutate(
        expires = as_date(expires, format = "%Y-%m-%d"),
        across(where(is.character), slack_emoji_to_unicode)
    )

# Keep the destination URL intact, but do not expose a long URL as the
# visible link label. Kramdown will not reliably wrap every long URL in the
# final Jekyll page, especially when it arrives through markdownify.
normalize_link_labels <- function(x, max_label_length = 80) {
    str_replace_all(
        x,
        str_c(
            "\\[(https?://[^\\]]{", max_label_length + 1,
            ",})\\]\\(([^)]+)\\)"
        ),
        "[View details](\\2)"
    )
}

old_data <- old_data |>
    mutate(links = normalize_link_labels(links))
```

## Slack Data

Next, we'll grab the Slack data from the Newsletter Content Triage
[List](https://usrse.slack.com/lists/T8ZT4PJSW/F08MLHYKALR). 
We could use the Slack API, but I think it's actually easier
to export the list content as CSV and `git` track it in the 
`_data` directory.

Importantly, the most recent file is the one we want to keep;
Slack doesn't automatically overwrite the CSV, so we have to manually delete old files.

```{r}
old_slack_files <- list.files(here("_data/newsletter"), pattern = "Newsletter_Content_Triage", full.names = TRUE) |>
    file.info() |>
    rownames_to_column("file") |>
    as_tibble() |>
    arrange(desc(mtime))

if(nrow(old_slack_files) > 1){
    slack_file <- old_slack_files |>
        slice(1)

    walk(old_slack_files$file[-1], fs::file_delete)
    fs::file_move(slack_file$file, here("_data/newsletter/Newsletter_Content_Triage.csv"))
} else {
    message("No old files to delete.")
}
```

```{r}

slack <- read_csv(here("_data/newsletter/Newsletter_Content_Triage.csv")) |>
    clean_names() |>
    mutate(across(where(is.character), slack_emoji_to_unicode))

section_col <- str_subset(names(slack), "what_section")

slack <- slack |>
    separate_rows(all_of(section_col), sep = ",\\s*") |>
    mutate(
        submission_date = as_date(submission_date, format = "%m/%d/%y, %I:%M %p"),
        staging = as.factor(staging)
        )    |>
    rename(section = all_of(section_col))
```

Now get the new events:

```{r}

fromJson_vec <- Vectorize(function(x){

    tryCatch({
        dat <- fromJSON(x)

        display <- slack_emoji_to_unicode(dat$displayName)
        url <- dat$originalUrl

        if (length(url) == 0 || is.na(url[[1]]) || !nzchar(url[[1]])) {
            return(x)
        }

        url <- as.character(url[[1]])
        display <- if (length(display) == 0) "" else as.character(display[[1]])

        if (is.na(display) || !nzchar(display) ||
            str_detect(display, "^https?://")) {
            display <- "View details"
        }

        str_c("[", display, "](", url, ")")
    }, error = function(e){
        x
    })
}, USE.NAMES = FALSE)

slack |>
    filter(str_detect(section, "Event")) |>
    filter(staging == "Approved" | staging == "Published") |>
    mutate(
        content_expiration_date = case_when(
            is.na(content_expiration_date) ~ as_date("2026-08-01"),
            TRUE ~ content_expiration_date
        ),
        event_date = case_when(
            is.na(event_date) ~ as_date("2026-08-01"),
            TRUE ~ event_date
        ),
        type = case_when(
            str_detect(str_to_lower(content), "submit|application|submission|deadline") ~ "opportunity",
            TRUE ~ "event"
        ),
        location = case_when(
            !is.na(location) ~ location,
            str_detect(str_to_lower(content), "online|zoom|webinar|virtual") ~ "Online",
            TRUE ~ "In Person"
        ),
        link = normalize_link_labels(fromJson_vec(link))
    ) -> events_opportunities
```

## Binding

Now we just put em together: 

```{r}
events_opportunities |>
    mutate(event_date = as.character(event_date)) |>
    select(
        expires = content_expiration_date,
        type = type,
        title = submission,
        when = event_date,
        where = location,
        preamble = content,
        links = link
    ) |>
    replace_na(list(when = as_date("2026-01-01"), where = "TBD", links = "TBD")) |>
    bind_rows(old_data) -> all_events_opportunities

all_events_opportunities <- all_events_opportunities |>
    mutate(
        across(where(is.character), slack_emoji_to_unicode),
        links = normalize_link_labels(links)
    )
```

Number of rows to publish next: 

```{r}
all_events_opportunities |>
    filter(expires >= Sys.Date()) |>
    dt2(compact = TRUE, striped = TRUE, hover = TRUE, font_scale = 0.85,
    options = list(pageLength = 10))
```

We can write this back to yaml now:

```{r}
all_events_opportunities |>
    filter(expires >= Sys.Date()) |>
    mutate(when = as.character(when)) |>
    mutate(expires = as.character(expires)) |>
    mutate(index= row_number()) |>
    group_by(index) |>
    group_split() |>
    write_yaml(here("_data/newsletter/events-opportunities.yml"))
```
